---
title: "R Notebook"
output: html_notebook
---
1.1 Local PostgreSQL connections

Q.1 Create a local PostgreSQL database called wrestling

Ans - Type in terminal psql : CREATE DATABASE wrestling;

Q.2 Seed the wrestling database from the file wrestling.sql using the psql PostgreSQL command line interface.
Ans - touch wrestling.sql

```{r}
library(keyring)

library(RPostgreSQL)

keyring_unlock("wrestling")

username <- key_get("wrestling_postgresql_username", keyring = "wrestling") 

password <- key_get("wrestling_postgresql_password", keyring = "wrestling")

keyring_lock("wrestling")

#db_connect <- dbConnect(drv = PostgreSQL(max.con = 2), user = username, password = password, dbname = "wrestling", host = "localhost")

my_db_connection <- dbConnect(drv = PostgreSQL(max.con = 2), user = username, password = password, dbname = 'wrestling', host = 'localhost')
```

Q.3 Using the R connection you just created, write and execute a SQL query to return all the wrestlers in the wrestlers table.

```{r}
dbGetQuery( conn = db_connect,
            "SELECT * FROM wrestlers")
```

Q.4 Explore the columns stored in the table. State which data type you think each column stores.

```{r}
data <- dbGetQuery( conn = db_connection,
                  statement = "SELECT * FROM employees")

class(data)
```




