---
title: "R Notebook"
output: html_notebook
---

Reading Data

The following data type is in excel format and hence we have used the read_excel function to read the .xls file,
Please note that the read_excel command only reads the first page of the data set provided, however there are a total of 4 sheers in the data set seabords.xls , to read the 2nd sheet we have to specify the sheet number, specifying the column type allows will allow the function to extract the data in a particular type i.e "skip","guess","date","logical","numeric","text" or as a "list".
So now we have the 4 excel sheets read as 4 vectors.
  1.ship_data_by_record_id
  2.bird_data_by_recrd_id
  3.ship_data_codes
  4.bird_data_codes

```{r}
library(tidyverse)
library(readxl)
install.packages("here")
library(here)

ship_data_by_record_id <- read_excel("data/seabirds.xls")

bird_data_by_recrd_id <- read_excel("data/seabirds.xls", col_types = "guess" ,  sheet = 2)

ship_data_codes <- read_excel("data/seabirds.xls",skip = 1, sheet = 3)

bird_data_codes <- read_excel("data/seabirds.xls",skip =1, sheet = 4)

```


#Cleaning the data

#Step 1: Viewing and analysing the data
```{r}
View(bird_data_by_recrd_id)
names(ship_data_by_record_id)
dim(ship_data_by_record_id)
head(ship_data_by_record_id)
tail(ship_data_by_record_id)
```

```{r}
library(janitor)

ship_data_by_record_id <- clean_names(ship_data_by_record_id)

```


Step 2: cleaning the column names and giving them a more meaningfull name for bird_data by looking at the bird data codes excel sheet...

```{r}

View(bird_data_by_recrd_id)
names(bird_data_by_recrd_id)
dim(bird_data_by_recrd_id)
head(bird_data_by_recrd_id)
tail(bird_data_by_recrd_id)

```

```{r}

bird_data_by_record_id <- clean_names(bird_data_by_recrd_id)

```



Step 2:

Dropping / Deleating columns which are not required , checking for repetations in data, missing data, and terating them as required.

```{r}

bird_data_by_record_id_dropped <- bird_data_by_record_id %>%
  select(record_id, bird_common_name,bird_scientific_name,bird_abbreviation, count)

---
title: "R Notebook"
output: html_notebook
---

Reading Data

The following data type is in excel format and hence we have used the read_excel function to read the .xls file,
Please note that the read_excel command only reads the first page of the data set provided, however there are a total of 4 sheers in the data set seabords.xls , to read the 2nd sheet we have to specify the sheet number, specifying the column type allows will allow the function to extract the data in a particular type i.e "skip","guess","date","logical","numeric","text" or as a "list".
So now we have the 4 excel sheets read as 4 vectors.
  1.ship_data_by_record_id
  2.bird_data_by_recrd_id
  3.ship_data_codes
  4.bird_data_codes

```{r}
library(tidyverse)

library(readxl)

library(openxlsx)

install.packages("here")

library(here)

ship_data_by_record_id <- read_excel("data/seabirds.xls")

ship_data_by_record_id

bird_data_by_recrd_id <- read_excel("data/seabirds.xls", col_types = "guess" ,  sheet = 2)

bird_data_by_recrd_id

ship_data_codes <- read_excel("data/seabirds.xls",skip = 1, sheet = 3)

ship_data_codes

bird_data_codes <- read_excel("data/seabirds.xls",skip =1, sheet = 4)

bird_data_codes
```


Cleaning the data

Step 1: cleaning the column names and giving them a more meaningfull name for ship_data
```{r}
View(bird_data_by_recrd_id)

names(ship_data_by_record_id)

dim(ship_data_by_record_id)

head(ship_data_by_record_id)

tail(ship_data_by_record_id)
```

```{r}
library(janitor)

names(ship_data_by_record_id) <- c("record", "record_id", "date", "time", "latitude", "longitude", " east_west_hemisphere", "ship_activity", " ship_speed", "ship_direction", "cloud_cover", "precipitation", "wind_speed", "wind_dicertion", "air_temperature", "sea_level_pressure", "sea_state", "sea_surface_temp", "sea_surface_salinity", "sea_floor_depth", "observer_name", "bird_count", "month", "sount_hemisphere_season","longitude_360", "latitude_cell_center", "longitude_cell_360")

ship_data_by_record_id <- clean_names(ship_data_by_record_id)

ship_data_by_record_id


```


Step 2: cleaning the column names and giving them a more meaningfull name for bird_data by looking at the bird data codes excel sheet...

```{r}

View(bird_data_by_recrd_id)

names(bird_data_by_recrd_id)

dim(bird_data_by_recrd_id)

head(bird_data_by_recrd_id)

tail(bird_data_by_recrd_id)

clean_names(bird_data_by_recrd_id)

```

```{r}
library(janitor)

names(bird_data_by_recrd_id) <- c("record", "record_id","bird_common_name", "bird_scientific_name", "bird_abbreviation", "age", "wandering_albertos_plumage_phase", "plumage_phase", "sex", "count", "feeding_num", "feeding_ocurance", "sitting_on_water_num", "sitting_on_water_ocur", " sitting on ice no", "stiiting on ice ocr", "sitting on ship ocur", "in hand ocur", "flying past num", "flying past ocr", "accompanying num", "accompanying occurances", "ship follow in wake num", "ship follow in wake occurances", "occurance of moult", "naturally feeding occurance")

bird_data_by_recrd_id

bird_data_by_record_id <- clean_names(bird_data_by_recrd_id)

bird_data_by_record_id

```



Step 2:

Dropping / Deleating columns which are not required , checking for repetations in data, missing data, and terating them as required.

```{r}

bird_data_by_record_id_cleaned <- bird_data_by_record_id %>%
  select(record_id, bird_common_name,bird_scientific_name,bird_abbreviation, count)

bird_data_by_record_id_cleaned

ship_data_by_record_id_cleaned <- ship_data_by_record_id %>%
  select(record_id, latitude, longitude, bird_count, month,sount_hemisphere_season )

ship_data_by_record_id_cleaned

```

Checking String Variables

Setting string to lower case across all

```{r}

bird_data_by_record_id_lower <- mutate_all(bird_data_by_record_id_cleaned, .funs=tolower)

bird_data_by_record_id_lower

ship_data_by_record_id_lower <- mutate_all(ship_data_by_record_id_cleaned, .funs = tolower)

ship_data_by_record_id_lower

```


Merging of the two tables 

```{r}
#scored_policies<-merge(x=policies,y=limits,by="State",all.x=TRUE) - source internet http://www.programmingr.com/tutorial/left-join-in-r/

bird_ship_data_record_id <- merge(x=bird_data_by_record_id_lower, y=ship_data_by_record_id_lower,by="record_id",all.x = TRUE)
  
bird_ship_data_record_id
```


Saving the table as a .csv file to the clean_data folder

```{r}
# Source https://datatofish.com/export-dataframe-to-csv-in-r/
# write.csv(Your DataFrame,"Path where you'd like to export the DataFrame\\File Name.csv", row.names = FALSE)

write_csv(bird_ship_data_record_id,"bird_ship_data_record_id.csv")
```

---
title: "R Notebook"
output: html_notebook
---

Reading Data

The following data type is in excel format and hence we have used the read_excel function to read the .xls file,
Please note that the read_excel command only reads the first page of the data set provided, however there are a total of 4 sheers in the data set seabords.xls , to read the 2nd sheet we have to specify the sheet number, specifying the column type allows will allow the function to extract the data in a particular type i.e "skip","guess","date","logical","numeric","text" or as a "list".
So now we have the 4 excel sheets read as 4 vectors.
  1.ship_data_by_record_id
  2.bird_data_by_recrd_id
  3.ship_data_codes
  4.bird_data_codes

```{r}
library(tidyverse)

library(readxl)

library(openxlsx)

install.packages("here")

library(here)

ship_data_by_record_id <- read_excel("data/seabirds.xls")
bird_data_by_recrd_id <- read_excel("data/seabirds.xls", col_types = "guess" ,  sheet = 2)
ship_data_codes <- read_excel("data/seabirds.xls",skip = 1, sheet = 3)
bird_data_codes <- read_excel("data/seabirds.xls",skip =1, sheet = 4)
```


#Cleaning the data

#Step 1: cleaning the column names and giving them a more meaningfull name for ship_data
```{r}
View(bird_data_by_recrd_id)

names(ship_data_by_record_id)

dim(ship_data_by_record_id)

head(ship_data_by_record_id)

tail(ship_data_by_record_id)
```

```{r}

names(ship_data_by_record_id) <- c("record", "record_id", "date", "time", "latitude", "longitude", " east_west_hemisphere", "ship_activity", " ship_speed", "ship_direction", "cloud_cover", "precipitation", "wind_speed", "wind_dicertion", "air_temperature", "sea_level_pressure", "sea_state", "sea_surface_temp", "sea_surface_salinity", "sea_floor_depth", "observer_name", "bird_count", "month", "sount_hemisphere_season","longitude_360", "latitude_cell_center", "longitude_cell_360")

ship_data_by_record_id <- clean_names(ship_data_by_record_id)
```


#Step 2: cleaning the column names and giving them a more meaningfull name for bird_data by looking at the bird data codes excel sheet...

```{r}

View(bird_data_by_recrd_id)

names(bird_data_by_recrd_id)

dim(bird_data_by_recrd_id)

head(bird_data_by_recrd_id)

tail(bird_data_by_recrd_id)

clean_names(bird_data_by_recrd_id)

```

```{r}
library(janitor)

names(bird_data_by_recrd_id) <- c("record", "record_id","bird_common_name", "bird_scientific_name", "bird_abbreviation", "age", "wandering_albertos_plumage_phase", "plumage_phase", "sex", "count", "feeding_num", "feeding_ocurance", "sitting_on_water_num", "sitting_on_water_ocur", " sitting on ice no", "stiiting on ice ocr", "sitting on ship ocur", "in hand ocur", "flying past num", "flying past ocr", "accompanying num", "accompanying occurances", "ship follow in wake num", "ship follow in wake occurances", "occurance of moult", "naturally feeding occurance")

bird_data_by_record_id <- clean_names(bird_data_by_recrd_id)
```


#Step 2:

#Dropping / Deleating columns which are not required , checking for repetations in data, missing data, and terating them as required.

```{r}

bird_data_by_record_id_dropped <- bird_data_by_record_id %>%
  select(record_id, bird_common_name,bird_scientific_name,bird_abbreviation, count)

ship_data_by_record_id_dropped <- ship_data_by_record_id %>%
  select(record_id, latitude, longitude, bird_count, month,sount_hemisphere_season )

```

#Checking String Variables

#Setting string to lower case across all

```{r}

bird_data_by_record_id_lower <- mutate_all(bird_data_by_record_id_dropped, .funs=tolower)

ship_data_by_record_id_lower <- mutate_all(ship_data_by_record_id_dropped, .funs = tolower)

```


#Merging of the two tables 

```{r}
#scored_policies<-merge(x=policies,y=limits,by="State",all.x=TRUE) - source internet http://www.programmingr.com/tutorial/left-join-in-r/

bird_ship_data_record_id <- merge(x=bird_data_by_record_id_lower, y=ship_data_by_record_id_lower,by="record_id",all.x = TRUE)
  
bird_ship_data_record_id
```


#Saving the table as a .csv file to the clean_data folder

```{r}
# Source https://datatofish.com/export-dataframe-to-csv-in-r/
# write.csv(Your DataFrame,"Path where you'd like to export the DataFrame\\File Name.csv", row.names = FALSE)

write_csv(bird_ship_data_record_id,"bird_ship_data_cleaned.csv")
```
