---
title: "R Notebook"
output: html_notebook
---

```{r}

primes <- c(2, 3, 5, 7, 11)

primes <- c(primes , 13, 17)

primes

```

TASK

```{r}


primes_intiger <- c(2L, 3L,5L, 7L, 11L, 13L, 17L)

primes_intiger
```

```{r}
primes_intiger * 10
```

```{r}
c(1, 2) + c(1, 2, 3, 4)
```

```{r}
c( "a" , "a", "b", "a")

as.factor(c("a" , "a", "b", "a"))
```

TASK

```{r}
fruit <- c("apple", "banaba", "cherry")

fruit <- c( fruit, "dragon fruit")

as.factor (fruit)

fruit
```
Task:
Create a sequence going from 10 to 100 in steps of 5. Name this vector answer_1.

Make a vector which repeats the number 8 six times. Name this vector answer_2.

Use rep() to make a vector like below. Name the vector answer_3.

## [1] 2 2 2 3 3 3 4 4 4
Make a vector by joining the answer_1, answer_2 and answer_3 vectors together.

What does sort() do to a logical vector?

Compare the sum of answer_1 with the sum of answer_1 reversed.

What happens when you call table() on a vector of unique elements.

What does sum() do to a logical vector?

```{r}
answer_1 <- seq( 10L, 100L, by = 5L )

answer_1

answer_2 <- rep( 8 , times = 6)

answer_2

answer_3 <- rep( c(2, 3, 4), each = 3)

answer_3

answer_4 <- (c(answer_1, answer_2, answer_3))

answer_4
```

TASK


For each of these questions, think about the most efficient way of finding the answer.

First manually calculate the answer to this sum, then convert it into R code and check that R gets the same answer as you.
3×2+22−5

Write R code to subtract 2 from each of the following numbers: 14, 18, 5.

Multiply the first number in this list by 1, the second by 2 and the third by 3: 5, 2, 3

For the following numbers, what is their remainder when divided by 5: 11, 17, 23.

```{r}
c( 3*2+ 2^2 -5)

c(14, 8, 5) - 2

c( 5, 2, 3) * c(1, 2 ,3)

c(11, 17, 23) %% 5
```
Task - 10 minutes

Prove that 1 is not equal to 2.
Does answer_3 contain a 5?
Make a vector that shows whether each value in answer_1 is divisible by 10.
How many elements of answer_1 are greater than or equal to 30?

```{r}
1 == 2

5 %in% answer_3

c(answer_1 %% 10 == 0)

sum (answer_1 >= 30 )
```
Task - 5 minutes

Select the third element of answer_1.
Select everything except the last element of answer_1.
Select all the elements of answer_1 which are less than 25.
```{r}
answer_1[3]

answer_1[- length(answer_1)]

answer_1[answer_1 < 25]
```

