---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```

```{r}
diamonds <- read_csv("diamonds.csv")
```

```{r}
View(diamonds)

```

```{r}
diamonds_temp <- diamonds %>%
  select("carat", "x", "y", "z")
```

```{r}
library(GGally)

ggpairs(diamonds_temp)
```
As we can see there is substancial correlation between these 4 variables,
Let us drop x , y and z and work only with carat hence forth in the data set
```{r}
diamonds <- diamonds %>%
  select(-c("X1", "x", "y", "z"))
         
View(diamonds)

```

Let us now create a regression model for the price of a diamond in terms of the possible predictor variables in the dataset.

Let us Use ggpairs() to investigate correlations between price and the predictors.
```{r}
ggpairs(diamonds)
```

Perform further ggplot visualisations of any significant correlations you find.
```{r}
diamonds %>%
  ggplot(aes(x = carat, y = price)) +
  geom_point()
```

```{r}
diamonds %>%
  ggplot(aes(x = cut , y = price)) +
  geom_boxplot()
```

```{r}
diamonds %>%
  ggplot(aes(x = color , y = price)) +
  geom_boxplot()
```

```{r}
diamonds %>%
  ggplot(aes(x = clarity , y = price)) +
  geom_boxplot()
```






Now we can try a regression fit using one or more of the categorical predictors cut, clarity and color, so let’s investigate these predictors:
Investigate the factor levels of these predictors.
Looks like we might have 
4 dummies for Cut    :Fair, Good, Very Good, Premium, Ideal 
6 dummies for Color  : D , E , F, G, H, I, J
7 dumies for Clarity :I1, SI2, SI1, VS2, VS1, VVS2, VVS1, IF

Please note in the outcome we will always loose 1 varaible column.

```{r}
diamonds_dummycut <- diamonds %>%
  fastDummies::dummy_cols(select_columns = "cut", remove_first_dummy = TRUE) 

diamonds_dummycut
```

```{r}
diamonds_dummycolor <- diamonds %>%
  fastDummies::dummy_cols(select_columns = "color", remove_first_dummy = TRUE) 

diamonds_dummycolor
```

```{r}
diamonds_dummyclarity <- diamonds %>%
  fastDummies::dummy_cols(select_columns = "clarity", remove_first_dummy = TRUE) 

diamonds_dummyclarity
```

Going forward we’ll let R handle dummy variable creation for categorical predictors in regression fitting (remember lm() will generate the correct numbers of dummy levels automatically, absorbing one of the levels into the intercept as a reference level)

Let us now start with simple linear regression.

Let us Regress price on carat and check the regression diagnostics.

```{r}
diamonds %>%
  ggplot(aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
model <- lm(price ~ carat, data = diamonds)
summary(model)
```
We can see that R-squared:  0.8493 (coefficient of determination) is almost close to 1 and we can also see p Value has got *** , hence we can say that its a good fit.

let us also take a look at the diagnostic plots

```{r}
plot(model)
```
Obesrvations :

Plots dont look that great, have concerns with Scale-Location as we can see a slight yet constant rise starting from 0.7 till the end

Also, Normal Q-Q graph has quiet a considerable number of deviations both top and bottom.

```{r}
model2 <- lm(price ~ carat + cut + clarity+ color, data = diamonds)

summary(model2)
```

