---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(infer)
library(janitor)
```

```{r}
ames <- read_csv("data/ames.csv") %>%
  clean_names()

glimpse(ames)
```
Plot a histogram of sale_price. Is the distribution normal?
```{r}
ames %>%
  ggplot(aes(x = sale_price)) +
  geom_histogram(col = "white")
```
Calculate the population parameter mean of sale_price. Store it in a variable, as we’ll need this value later when we check if each CI contains the parameter value or not.
```{r}
mean_ames <- mean(ames$sale_price)

mean_ames
```
take 1000 random samples each of size 100 from sale_price in ames
```{r}
rep_sample_1000 <- ames %>%
  rep_sample_n(size = 100, reps = 1000) %>%
  group_by(replicate) %>%
  summarise(
    mean_sale_price = mean(sale_price)
  )


rep_sample_1000 %>%
  ggplot(aes(x = mean_sale_price)) +
  geom_histogram(col = "white")
```
Calculate SD of rep_sample_1000
```{r}
rep_sample_1000 %>%
  summarise(stderr = sd(mean_sale_price))
```
2. create a boostrap sampling distribution of mean(sale_price) from each 100-observation sample (use 1000 reps)

Step : 1 ( First create a single_sample out of the 1000 samples we have created)
```{r}
single_sample_1000 <- ames %>%
  rep_sample_n(size = 1000, reps = 1) %>%
  ungroup() %>%
  select(sale_price)

single_sample_1000

```
Generate bootstrap distribution
```{r}
bootstrap_rep_sample <- single_sample_1000 %>%
  specify(response = sale_price) %>%
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "mean")

bootstrap_rep_sample
```
3. calculate the 95% CI for each bootstrap distribution
```{r}
bootstrap_tenure_ci <- bootstrap_rep_sample %>%
  get_confidence_interval(level = 0.95, type = "percentile")
bootstrap_tenure_ci
```

