---
title: "R Notebook"
output: html_notebook
---

Reading Data

The following data type is in excel format and hence we have used the read_excel function to read the .xls file,
Please note that the read_excel command only reads the first page of the data set provided, however there are a total of 4 sheers in the data set seabords.xls , to read the 2nd sheet we have to specify the sheet number, specifying the column type allows will allow the function to extract the data in a particular type i.e "skip","guess","date","logical","numeric","text" or as a "list".
So now we have the 4 excel sheets read as 4 vectors.
  1.ship_data_by_record_id
  2.bird_data_by_recrd_id
  3.ship_data_codes
  4.bird_data_codes

```{r}
library(tidyverse)

library(readxl)

library(openxlsx)

ship_data_by_record_id <- read_excel("data/seabirds.xls")

ship_data_by_record_id

bird_data_by_recrd_id <- read_excel("data/seabirds.xls", col_types = "guess" ,  sheet = 2)

bird_data_by_recrd_id

ship_data_codes <- read_excel("data/seabirds.xls",skip = 1, sheet = 3)

ship_data_codes

bird_data_codes <- read_excel("data/seabirds.xls",skip =1, sheet = 4)

bird_data_codes
```


Cleaning the data

Step 1: cleaning the column names and giving them a more meaningfull name.
```{r}
View(ship_data_by_record_id)

names(ship_data_by_record_id)

dim(ship_data_by_record_id)

head(ship_data_by_record_id)

tail(ship_data_by_record_id)
```

```{r}
library(janitor)

names(ship_data_by_record_id) <- c("record", "record_id", "date", "time", "latitude", "longitude", " east_west_hemisphere", "ship_activity", " ship_speed", "ship_direction", "cloud_cover", "precipitation", "wind_speed", "wind_dicertion", "air_temperature", "sea_level_pressure", "sea_state", "sea_surface_temp", "sea_surface_salinity", "sea_floor_depth", "observer_name", "bird_count", "month", "sount_hemisphere_season","longitude_360", "latitude_cell_center", "longitude_cell_360")

ship_data_by_record_id <- clean_names(ship_data_by_record_id)

ship_data_by_record_id


```

Step 2:

Dropping / Deleating columns which are not required , checking for repetations in data, missing data, and terating them as required.

