---
title: "R Notebook"
output: html_notebook
---
1.Read in the state_income_data.csv file using the appropriate function.

```{r}
read.csv("data/state_income_data.csv")
```

Q2.How do you specify your own column names when using read_csv()?

Ans - For the given data set state_income_data.csv

read_csv("data/state_income_data.csv",
  col_names = c(
    "Column 1",
    "Column 2",
    "Column 3",
    "Column 4"
  )
  
  
Q3.Read in the cars.csv file and use the col_types() function to ensure you load in:

X1 as character
mpg, cyl, disp, hp as double
vs as logical

```{r}
read.csv("data/cars.csv")
```

```{r}

library(readr)
cars_chnaged_datatype <- read_csv(("data/cars.csv"), col_types = cols(
                  X1   = col_character(),
                  mpg  = col_double(),
                  cyl  = col_double(),
                  disp = col_double(),
                  hp   = col_double(),
                  vs   = col_logical()
                  ))

cars_chnaged_datatype
```

Q.4 - Read in the scotland.txt file using the appropriate function.

```{r}

library(readr)
scotland_data <- read_delim("data/scotland.txt", delim = ";" , skip = 1)

scotland_data
```

Q.5 How do you read text into a one element character vector?

Ans : Sometimes we want to read in plain text files that are not delimited. We can read in a text file using read_file. The result will be stored as character vector of length one.

Q.6 Read in the steam-200k-sample.csv file. Fix any problems you may encounter.

Please check with instructor

```{r}

steam_200k_sample <- read_file("data/steam-200k-sample.csv", skip = 4)

steam_200k_sample

```

Q.7 Read in the historic_eu_payments.tsv file.

```{r}
library(readxl)

historic_eu_payments <- read_tsv("data/historic_eu_payments.tsv")

historic_eu_payments
```





Q.8 How do you read lines of text into a multi-element text vector?

Ans : By using the read_lines() function






Q.9 Read in the JSON file from the following url: http://makemeanalyst.com/wp-content/uploads/2017/06/sample_json.txt

```{r}
library(jsonlite)
library(readr)

jason_from_url <- read_file("http://makemeanalyst.com/wp-content/uploads/2017/06/sample_json.txt")

jason_from_url
```

