---
title: "R Notebook"
output: html_notebook
---
2.2.1 
Write a function which multiples all numbers by 15 and adds 4.
```{r}
function_mul_15_add_4 <- function (x) {
  return(x * 15 + 4 )
}
```


```{r}
function_mul_15_add_4(2)
                      
```

2.2.2
Write a function that returns “bird”, if the input is “goose”, “duck” or “pigeon”, and returns “not bird” for all other input.
```{r}

function_bird_notbird <- function(x) {
  
  if (x %in% c("goose", "duck", "pigeon")) {
    
    return ("bird") 

  } else {
    
    return ("not bird")}
  
}
```

```{r}
function_bird_notbird("goose")
```

Using Only If statement
```{r}
function_bird_notbird_if_only <- function(x) { 
  
if ( x %in% c("goose", "duck", "pigeon")){
  
    print("bird") }
if ( !x %in% c("goose", "duck", "pigeon"))
  {
  print("not bird")
  }  
}

function_bird_notbird_if_only("Amit")
```


Write a function which multiplies all numbers by 10. If the input isn’t a number, the function should return “Invalid input”.

Note : Ask Instructur Why do we get an eror message when we do not type words in " "

```{r}
function_multiply_by_10 <- function(x){
  
  if (is.numeric(x)) {
    
    return ( x * 10)}
  
  if (!is.numeric(x)) {
    
    return ("Invalid Input")
  }
}

function_multiply_by_10("Amit")
```

2.3 ifelse(), recode() and case_when()

For the vector below, make a new vector which is equal to “Big” every time the element of sizes is greater than 10, and equal to “Small” every time the element of sizes is less than or equal to 10.

```{r}
sizes <- c(13, 15, 10, 8, 22)

ifelse(sizes > 10, "Big", "Small")

  

```


2.3.2
For the vector below convert:
apple → alpaca
banana → bonobo
cherry → chipmunk

```{r}
fruit <- c("apple", "apple", "banana", "cherry", "cherry")

library(dplyr)

dplyr::recode(fruit, "apple" = "alpaco", "banana" = "bonobo", "cherry" = "chipmunk")


```

2.3.3
The two vectors below show the ages at marriage of each member of a couple. The first elements of each vector correspond to the first couple, the second elements to the second couple, and so on.

Create a new vector which is equal to:

“Wife older” whenever the wife’s age is highest “Husband older” when the husband’s age is highest “Same age” when they are the same age

```{r}
wife_age <- as.factor(c(29L, 30L, 36L, 31L, 22L))
husband_age <- as.factor(c(28L, 35L, 32L, 31L, 29L))

wife_husband_age_comparison <- 
  
case_when(
  wife_age > husband_age ~ "Wife Older",
  wife_age = husband_age ~ "Same Age",
  wife_age < husband_age ~ "Husband Older"
)


```

2.4 Loops

```{r}
my_list <- list(
  4,
  "apple",
  10.5,
  c("is", "an", "apple"),
  "apple",
  c("i", "like", "apples"),
  8
)

my_list
```

2.4.1
Use a loop to print out every element of my_list which is of type character. Decide if you need a for loop or a while loop.

```{r}
for ( x in my_list) {
  if ( x == "character") {
    print( x )
  }
}
```

```{r}
product_matrix_12 <- matrix(nrow = 12, ncol = 12)
                         
num_rows <- dim(product_matrix_12)[1]

num_cols <- dim(product_matrix_12)[2]
  
for(row_index in 1:num_rows){  
  
  for(col_index in 1:num_cols){ 
    
    product_matrix_12[row_index, col_index] <- row_index * col_index    
  }
}

product_matrix_12
```

3.2
Write a function that takes a vector, and creates a text art histogram, like the one below.

You may want to use the function cat(). This is similar to print(), but prints directly, without vector markers. You can also include newlines "\n" when using cat().

```{r}
x <- c ( 3, 1, 7, 4, 2)

cat("iteration =  ", x , "\n")

x
  
```

3.4
By considering the terms in the Fibonacci sequence whose values do not exceed four million, find the sum of the even-valued terms.

You may want to read up on the definition of the Fibonacci sequence.

Example : 1, 2, 3, 5, 8, 13, 21, 34, 55, 89

```{r}
#Fibonacci_seq_4_million <- function(x){
  
#  first_number <- 1
#  seconce_number <- first_number + first_number
#  third_number <- 
#}

multiply_by_3_add_2 <- function(x){
  
  times_3 <- x*3
  add_2 <- times_3 + 2
  
  return(c(times_3 = times_3, add_2 = add_2))
}

multiply_by_3_add_2(10)
```

