---
title: "R Notebook"
output: html_notebook
---

Setting up Required Libraries and reading the CSV files.
```{r}
library(tidyverse)
library(readr)
library(here)
here::here()
decathlon <- read_csv(here("clean_data/decathlon_clean.csv"))
View(decathlon)
names(decathlon)
```

Who had the longest long jump seen in the data?
```{r}
  decathlon %>%
  select(player_name,competition,long_jump) %>%
  arrange(desc(long_jump)) %>%
  head(1)
```

What was the average 100m time in each competition?
```{r}
  decathlon %>%
  select(player_name,x100m,competition) %>%
  group_by(competition)%>%
  summarise(mean(x100m))
```

Who had the highest total points across both competitions?
```{r}
decathlon %>%
  distinct(player_name, competition, points) %>%
  group_by(player_name) %>%
  summarise(sum_of_points = sum(points)) %>%
  arrange(desc(sum_of_points)) %>%
  head(1)
```

What was the shot-put scores for the top three competitors in each competition?
```{r}

  filter(decathlon, competition == "Decastar")%>%
  select(player_name,shot_put,points,competition)%>%
  group_by(competition)%>%
  arrange(desc(points))%>%
  head(3)

  filter(decathlon, competition == "OlympicG")%>%
  select(player_name,shot_put,points,competition)%>%
  group_by(competition)%>%
  arrange(desc(points))%>%
  head(3)
  
```

What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}
  filter(decathlon, x400m > 50) %>%
  select(player_name, x400m, points) %>%
  summarise(avarage_points_who_ran_400m_in_more_than_50sec = mean(points))

  filter(decathlon, x400m < 50) %>%
  select(player_name, x400m, points) %>%
  summarise(avarage_points_who_ran_400m_in_less_than_50sec = mean(points))
```



