---
title: "R Notebook"
output: html_notebook
---
Now we’ll go back to CI creation in the normal fashion. We’ll take the ames data from the CIs lab earlier today and regard it now as a sample, we won’t be drawing any smaller samples from within it. This is the usual situation in an analysis: you use all the data available to you!



1.Load the data again, clean_names(), and re-familiarise yourself with it

```{r}
library(tidyverse)
library(infer)
library(janitor)
```
reading the file and cleaning the names
```{r}
ames <- read_csv("data/ames.csv")

ames <- clean_names(ames)

View(ames)
names(ames)

```


2. Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?
Ans : Plot a histogram of lot_area. Is the distribution normal? Also calculate the mean
```{r}
mean(ames$lot_area)
```

```{r}
ames %>%
  ggplot(aes(x = lot_area)) +
  geom_histogram(col = "white")
```
It is not normal, It looks Right Skewed

3.Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.

```{r}
bootstrap_distribution <- ames %>%
  specify(response = lot_area) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")

bootstrap_distribution %>%
  visualise(bins = 30)
```

4. Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution

```{r}
bootstrap_ci <- bootstrap_distribution %>%
  get_confidence_interval(level = 0.95, type = "percentile")
bootstrap_ci
```
step b : now visualising them
```{r}
bootstrap_distribution %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = bootstrap_ci)
```


You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?

```{r}
bootstrap_ci_99 <- bootstrap_distribution %>%
  get_confidence_interval(level = 0.99, type = "percentile")
bootstrap_ci_99
```
now lets visualise it
```{r}
bootstrap_distribution %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = bootstrap_ci_99)
```

Calculate the point estimate of the mean(lot_area)
```{r}
bootstrap_distribution %>%
  summarise(mean(stat))
```


