---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(lubridate)
library(tsibble)
library(tsibbledata)
library(fable)
```

1 Question 1
Load in the nyc_bikes data from the tsibbledata package. Have an initial look at it to see what you’re working with. Create three new columns: one that stores only the year the bike was used, one that stores only the month the bike was used, and one that stores the date. Use the data stored in start_time to create these new columns.

```{r}
nyc_bikes <- nyc_bikes %>%
  mutate( date = as_date(start_time), year = year(start_time), month = month(start_time))

nyc_bikes

```

2 Question 2
Summarise the number of bike hire counts by month. Make a plot of this data. *Hint: remember that to group time series (tsibble) data, you need to use index_by instead of group_by before your summarise function. What does this plot tell you about the time series? Do you think this downsampled data would be adequate to build a forecast with?

```{r}

nyc_bikes %>%
  index_by(month) %>%
  summarise(Sum = n()) %>%
  ggplot() + 
  geom_line(aes(x = month, y = Sum), colour = "blue")

```

3 Question 3
Now Summarise the number of bike hire counts by date. Make a plot of this new aggregated data. What does this plot tell you about the time series? Would this data be preferrable for time series forecasting compared to the monthly data?

```{r}
nyc_bikes_data_summary <- nyc_bikes %>%
  index_by(date) %>%
  summarise(Sum = n())


  ggplot(nyc_bikes_data_summary) + 
  geom_line(aes(x = date, y = Sum), colour = "blue")

nyc_bikes_data_summary
```

Question 4:

Let’s begin to build a model. For this, we will use the downsampled by date dataset we created above in question 3. 

```{r}
# fill the gaps : here i'll use median as it's a decent summary of how bike rentals work (not overly affected by seasonal changes)

nyc_bikes_filled <- nyc_bikes_data_summary %>%
  fill_gaps(Sum = as.integer(median(Sum)))

ggplot(nyc_bikes_filled) + 
  geom_line(aes(x = date, y = Sum), colour = "blue")

nyc_bikes_filled
```





Let’s begin to build a model. For this, we will test the NAIVE, MEAN, and SNAIVE model.

```{r}
fit <- nyc_bikes_filled %>%
  model(
    snaive = SNAIVE(Sum),
    mean_model = MEAN(Sum),
    arima = ARIMA(Sum)
  )

fit
```

5 Question 5
Now we have our model fit, build a forecast to predict bike use over the next four months. Plot your models alongside your data.
Hint: forecast parameter would be roughly 120 (30 days x 4 months)

```{r}
forecast_1 <- fit %>%
  fabletools::forecast(h = 120)
forecast_1

#Plotting our forecast

forecast_1 %>%
  autoplot(nyc_bikes_filled, level = NULL) +
  ggtitle("Forecasts for NYC Cycles") +
  xlab("Month") +
  guides(colour = guide_legend(title = "Forecast"))
```

6 Question 6
Test your model accuracy : choose a training data set from your main dataset, build a forecast on the training set, and then plot the training set forecast against the real data. Calculate model accuracy.

```{r}
# Set training data from 2018-Jan to 2018-Sept
train <- nyc_bikes_filled %>%
  filter_index("2018-01" ~ "2018-09")

# run the model on the training set 
fit_test <- train %>%
  model(
    mean_model = MEAN(Sum),
    arima = ARIMA(Sum),
    snaive = SNAIVE(Sum))
```

Now let’s generate a forecast for the last quatrer Q4, from our training set data, and then plot it against the actual values.

```{r}
# forecast from the training set
forecast_test <- fit_test %>% 
  fabletools::forecast(h = 90)

# Plot forecasts against actual values
forecast_test %>%
  autoplot(train, level = NULL) +
    autolayer(filter_index(tsibbledata::nyc_bikes, "2018-10" ~ .), color = "black") +
    ggtitle("Forecasts for quarterly NYC Cycles") +
    xlab("Month") + ylab("Number of Bikes hired") +
    guides(colour=guide_legend(title="Forecast"))
```

8 Question 8
Make a simple ggplot (geom_point) which plots the start longitude and latitudes of each bike. Create a separate facet for each bike_id. Colour the dots in by month of use. What does this tell you about what month each bike was used most in?

Do the same for the end longitude and latitudes.

```{r}
library(sf)
```












