---
title: "R Notebook"
output: html_notebook
---
```{r}
library(readr)
library(stringr)
library(tidyverse)

code_clan_tweets <- read_csv("data/code_clan_tweets.csv")

nrow(code_clan_tweets)

ncol(code_clan_tweets)

dim(code_clan_tweets)

names(code_clan_tweets)

head (code_clan_tweets)

tail(code_clan_tweets)
```

```{r}
hist(code_clan_tweets$favorite_count)

boxplot(code_clan_tweets$favorite_count)

summary(code_clan_tweets$favorite_count)
```



3.Find the total number of favourited tweets (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
favourite_count_including_quote <- select(code_clan_tweets, is_quote, favorite_count)

count(favourite_count_including_quote)
```

```{r}

total_favourite_tweet_excluding_quotes <- filter(favourite_count_including_quote, is_quote == "FALSE")  
                                           count(favourite_count_excluding_quotes) %>%

```

```{r}
total_favourite_tweet_excluding_quotes
```

Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.


```{r}
data_omoying_quoted_tweets <- filter(code_clan_tweets, is_quote =="False")
                               
mean(code_clan_tweets$retweet_count)
                              
group_by(data_omoying_quoted_tweets)

data_omoying_quoted_tweets



```


Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.
```{r}



```


6.Find the mean number of characters that a CodeClan tweet contains
```{r}
mean(code_clan_tweets$display_text_width)
```

7.The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
code_clan_info <- read_csv("data/code_clan_info.csv")
```

```{r}
code_clan_info

nrow(code_clan_info)

ncol(code_clan_info)

dim(code_clan_info)

names(code_clan_info)

head (code_clan_info)
```

```{r}
code_clan_tweets_info <- full_join(code_clan_tweets, code_clan_info, by ="tweet_id")

code_clan_tweets_info
```


8.From your new joined data, create a new tibble codeclan_hashtags containing only the tweet_id and hashtags in lowercase for analysis. Keep only those tweets with hashtags
```{r}

# step one by using the mutate function,change all the text to lower case,
# step two select only the tweets which have the "# "spevial character in them


codeclan_hashtags <- select(code_clan_tweets_info$tweet_id,"text == # ")


```

